Symmetric Circuits for Rank Gates

### 2017-07-18 paper.tex
asdfasdf

\usepackage[margin=1in]{geometry} % set the margins to 1in on all sides
\usepackage{graphicx} % to include figures
\usepackage{amsmath} % great math stuff
\usepackage{amsfonts} % for blackboard bold, etc
\usepackage{amsthm} % better theorem environments

@@

\newtheorem{thm}{Theorem} \newtheorem{claim}{Claim}
\newtheorem{remark}{Remark} \newtheorem{definition}{Definition}
\newtheorem{lem}[thm]{Lemma} \newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary} \newtheorem{conj}[thm]{Conjecture}

@@

\newcommand{\bd}[1]{\mathbf{#1}} % for bolding symbols
\newcommand{\RR}{\mathbb{R}} % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}} % for Integers
\newcommand{\MB}{\mathbb{B}_{\matsym}} % for Integers
\newcommand{\SB}{\mathbb{B}_{\sym}} % for Integers
\newcommand{\col}[1]{\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}
\newcommand{\stab}{\text{\textbf{Stab}}}
\newcommand{\setstab}{\text{\textbf{SetStab}}}
\newcommand{\matstab}{\text{\textbf{MatStab}}}
\newcommand{\inv}{\text{\textbf{inv}}}
\newcommand{\aut}{\text{\textbf{Aut}}}
\newcommand{\kernal}{\text{\textbf{ker}}}
\newcommand{\alt}{\text{\textbf{Alt}}}
\newcommand{\sym}{\text{\textbf{Sym}}}
\newcommand{\rank}{\text{\textbf{rk}}}
\newcommand{\orb}{\text{\textbf{Orb}}}
\newcommand{\SP}{\text{\textbf{SP}}}
\newcommand{\maj}{\text{\textbf{Maj}}}
\newcommand{\dom}{\text{\textbf{Dom}}}
\newcommand{\child}{\text{child}}
\newcommand{\countgate}{\text{count}}
\newcommand{\agree}{\text{AGREE}}
\newcommand{\supp}{\text{SUPP}}
\newcommand{\rankparam}{\text{PARAM}}

@@

\nocite{*}

\title{Circuits}

\author{Gregory Wilsenach \\ 
Computer Laboratory \\
University of Cambridge}

@@

\chapter{Background}

\chapter{Symmetric Circuits}

\chapter{Extending Symmetric Circuits}

\chapter{Matrix-Symmetric Circuits}

The symmetric circuits of Anderson and Dawar \cite{} have been shown to express
$FPC$.

In this section we develop a notion of a matrix-symmetric circuit, an extension
of the symmetric circuit defined by Anderson and Dawar \cite{}. This is my
latest version of the support and circuit capturing result.

\section{}

\section{Circuits and Symmetry}
\begin{definition}
  Let $A$ and $B$ be sets, say that a function $f: \{ 0,1 \}^{A \times B}
  \rightarrow \{ 0,1 \}$ is matrix-symmetric if for any $\omega: A \times B
  \rightarrow \{0, 1\}$ and $(\alpha, \beta) \in \sym_A \times \sym_B$ we have
  $f((\alpha, \beta)\cdot \omega) = f (\omega)$.
  \\
  A function $f: \{0,1\}^* \rightarrow \{0,1\}$ is matrix-symmetric iff for any
  two finite sets $A$, $B$ we have that that $f|_{A \times B}$ is
  matrix-symmetric (where $f|_{A \times B}$ is the restriction of the function
  $f$ to inputs sequences of size $\vert A \times B \vert$ and where elements of
  the input sequence are labelled by elements of $A \times B$).
\end{definition}

\section{The Support Theorem}

Without a loss of generality we assume that $A$ and $B$ are always initial
segments of the natural numbers.

% Let $C$ be a circuit over some Boolean basis. For each gate $g \in C$ we
% associate a relation $\omega_g$ (or just $\omega$ when the context makes it
% clear). We say a gate is symmetric if that relation

% So far we have defined a circuit over a Boolean basis where the circuit
% elements all compute symmetric functions. The definition of an induced
% automorphism only makes. As such, we begin by augmenting the definition of a
% circuit and automorphism.

\begin{definition}[Circuits on Structures]
  For $\mathbb{B}_{\sym}$ a basis of Boolean symmetric functions,
  $\mathbb{B}_{\matsym}$ a basis of Boolean matrix-symmetric functions and
  $\tau$ a set of relation symbols, we define a $(\mathbb{B}_\sym,
  \mathbb{B}_\sym, \tau)$-circuit $C_n$ computing a $q$-ary query $Q$ is a
  structure $\langle G, W, \Omega, \Sigma, \Lambda, L\rangle$.
  \begin{itemize}
  \item $G$ is called the set of gates of $C_n$ and $\vert C_n \vert := \vert G
    \vert$.
  \item $W \subseteq G \times G$, where $W$ is called the wires of the circuit.
    $(G,W)$ must be a directed acyclic graph. For $g \in G$ we $H_g := \{ h \in
    C_n : W(h,g)\}$ be the set of children of $g$.
  \item $\Omega$ is an injective function from $[n]^q$ to $G$. The gates in the
    image of $\Omega$ are called the output gates. When $q = 0$, $\Omega$ is a
    constant function mapping to a single output gate.
  \item $\Sigma$ is a function from $G$ to $\mathbb{B}_\sym \uplus
    \mathbb{B}_\matstab \uplus \tau \uplus \{0,1\} $ which maps input gates to
    $\tau \uplus \{0,1\}$ and where $\Sigma^{-1} (0) \leq 1$ $\Sigma^{-1} (1)
    \leq 1$ and the internal gates get mapped into $\mathbb{B}_\sym \uplus
    \mathbb{B}_\matstab$. Gates mapped to $\tau$ are called relational gates and
    gates mapped to 1 or 0 are called constant gates.
  \item $\Lambda$ is a sequence of injective functions $(\Lambda_R)_{R \in
      \tau}$ where for each $R \in \tau$, $\Lambda_R$ maps each relational gate
    $g$ with $R = \Sigma (g)$ to the tuple $\Lambda_R (g) \in [n]^r$, where $r$
    is the arity of the symbol $R$. When no ambiguity arises we write $\Lambda
    (g)$ for $\Lambda_R (g)$.
  \item $L$ is the set of labelings of $C_n$. Let $g \in G$ be an internal gate
    and $H$ be the children of $g$. Then let $S$ be a set where $\{0,1\}^S =
    \dom(\Sigma (g))$. Then $L$ is a function that assigns to $g$ an onto
    function $\omega_g:S \rightarrow H$. We call this the matrix labelling of
    $g$.
  \end{itemize}
\end{definition}

% \begin{definition}
%   let $C = \langle G, W, \Omega, \Sigma, \Lambda, L\rangle$ be a
%   $(\mathbb{B}_\sym, \mathbb{B}_\sym, \tau)$-circuit, then take $g_1,g_2 \in
%   G$ such that $\Sigma (g_1) \Sigma(g_2) \in \mathbb{B}_\matsym$.

\begin{definition}
  Let $\omega_1: A \times B \rightarrow H$ and $\omega_2: A' \times B'
  \rightarrow H'$ define the equivalence relation $\sim$ as $\omega_1 \sim
  \omega_2$ iff $\omega_1$ and $\omega_2$ have the same domain and co-domain and
  there exists $(\alpha, \beta) \in \sym_A \times \sym_B$ and for all $(a,b) \in
  A \times B$ we have $\omega_1 = \omega_2 \circ (\alpha, \beta)$.
\end{definition}

% \begin{definition}
%   Let $\omega_1: A \rightarrow H$ and $omega_2:A' \rightarrow H'$. Define the
%   equivalence relation $\sim$ as $\omega_1 \sim \omega_2$ iff $\omega_1$ and
%   $\omega_2$ have the same domain and co-domain, and
% \end{definition}

\begin{definition}[Automorphism]
  let $C = \langle G, W, \Omega, \Sigma, \Lambda, L\rangle$ be a
  $(\mathbb{B}_\sym, \mathbb{B}_\sym, \tau)$-circuit computing at $q$-ary query
  on structures of size $n$. Let $\sigma \in \sym_n$ and $\pi: G \rightarrow G$
  be a bijection such that
  \begin{itemize}
  \item for all gates $g, h \in G$, $W(g,h)$ iff $W(\pi g, \pi h)$,
  \item for all output tuples $x \in [n]^q$, $\pi \Omega (x) = \Omega (\sigma
    x)$,
  \item for all gates $g \in G$, $\Sigma (g) = \Sigma (\pi g)$, and
  \item for each relational gate $g \in G$, $\sigma \Lambda (g) = \Lambda (\pi
    g)$
  \item for each internal gate $g$ if $\Sigma (g) \in \mathbb(B)_\matsym$. Then
    we have that $L(\pi g) \sim \pi \circ L(g)$.
  \end{itemize}
\end{definition}

\begin{remark}
  I also need a new notion of robustness here in order to collapse gates that
  have the same children and equivalent labelings are properly regarded as being
  the same gate.
\end{remark}
% \begin{definition}
%   Let $C$ be a circuit, $g$ be a gate in $C$, $H$ be $g$'s set of children and
%   let $\omega: A \times B \rightarrow H$ be a surjection, where $A$ and $B$
%   are sets and $g$ takes $\vert A \vert \cdot \vert B \vert$ inputs. Call such
%   an $\omega$ a matrix labeling for $g$.
% \end{definition}

% \begin{definition}
%   Let $C$ be a circuit and $g_1$ and $g_2$ be gates in $C$. Let $\sigma \in
%   \sym_n$ be a permutation such
% \end{definition}

>>>>

\usepackage[margin=1in]{geometry} % set the margins to 1in on all sides
\usepackage{graphicx} % to include figures
\usepackage{amsmath} % great math stuff
\usepackage{amsfonts} % for blackboard bold, etc
\usepackage{amsthm} % better theorem environments
\usepackage{xspace}

@@

\newtheorem{thm}{Theorem} \newtheorem{claim}[thm]{Claim}
\newtheorem{remark}[thm]{Remark} \newtheorem{definition}[thm]{Definition}
\newtheorem{lem}[thm]{Lemma} \newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary} \newtheorem{conj}[thm]{Conjecture}

@@

\newcommand{\reals}{\mathbb{R}}
\newcommand{\nats}{\mathbb{N}}

\newcommand{\logic}[1]{\text{\upshape #1}\xspace}
\newcommand{\FPC}{\logic{FPC}}
\newcommand{\FO}{\logic{FO}}
\newcommand{\FP}{\logic{FP}}
\newcommand{\FPR}{\logic{FPrk}}
\newcommand{\CPTC}{\logic{$\tilde{\text C}$PT(Card)}}

\newcommand{\PT}{\ensuremath{\mathrm{P}}\xspace}

\newcommand{\bd}[1]{\mathbf{#1}} % for bolding symbols
\newcommand{\RR}{\mathbb{R}} % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}} % for Integers
\newcommand{\MB}{\mathbb{B}_{\matsym}} % for Integers
\newcommand{\SB}{\mathbb{B}_{\sym}} % for Integers
\newcommand{\col}[1]{\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}
\newcommand{\stab}{\text{\textbf{Stab}}}
\newcommand{\setstab}{\text{\textbf{SetStab}}}
\newcommand{\matstab}{\text{\textbf{MatStab}}}
\newcommand{\inv}{\text{\textbf{inv}}}
\newcommand{\aut}{\text{\textbf{Aut}}}
\newcommand{\kernal}{\text{\textbf{ker}}}
\newcommand{\alt}{\text{\textbf{Alt}}}
\newcommand{\sym}{\text{\textbf{Sym}}}
\newcommand{\std}{\text{\textbf{std}}}
\newcommand{\rank}{\text{\textbf{rk}}}
\newcommand{\orb}{\text{\textbf{Orb}}}
\newcommand{\SP}{\text{\textbf{SP}}}
\newcommand{\SPs}{\text{\textbf{SPs}}}
\newcommand{\maj}{\text{\textbf{maj}}}
\newcommand{\dom}{\text{\textbf{Dom}}}
\newcommand{\img}{\text{\textbf{Img}}}
\newcommand{\child}{\text{child}}
\newcommand{\countgate}{\text{count}}
\newcommand{\agree}{\text{AGREE}}
\newcommand{\supp}{\text{SUPP}}
\newcommand{\rankparam}{\text{PARAM}}
\newcommand{\consp}{\text{sp}}
\newcommand{\consps}{\text{sps}}
\newcommand{\row}{\text{row}}
\newcommand{\column}{\text{col}}
\newcommand{\rowsp}{\text{r}}
\newcommand{\colsp}{\text{c}}
\newcommand{\EV}{\text{EV}}
\newcommand{\type}{\text{type}}
\newcommand{\types}{\text{types}}
\newcommand{\spstab}[1]{\stab (\consp (#1))}

@@

\title{Symmetric Circuits for Rank Gates}

\author{Anuj Dawar and Gregory Wilsenach \\ 
Computer Laboratory \\
University of Cambridge}

@@

A finite relational structure, such as a graph, can be represented as a binary
string in a standard way (for instance, by enumerating its adjacency matrix).
When we speak of the computational complexity of properties of graphs, we assume
such an encoding in order to present the structure as input to a machine or
circuit or other appropriate computational model. The properties of structures
we are interested in studying are invariant under isomorphisms. This requirement
translates into an invariance condition on sets of strings---say that a language
$L \subseteq \{0,1\}^\star$ is a \emph{graph property} if, whenever $x \in L$
for a string $x$ representing a graph $G$, we have $y\in L$ for any string $y$
obtained from $x$ by re-ordering the vertices of $G$. The central open question
in descriptive complexity---whether there is a logical characterisation of the
polynomial-time decidable properties of finite relational structures--- can then
be understood as the question of characterizing those languages $L$ that are in
$\PT$ and are graph properties.

Anderson and Dawar~\cite{AndersonD17} study this question through the lens of circuit
complexity, by considering \emph{symmetric circuits}. These are circuits that
necessarily accept graph properties, with this fact explicitly witnessed by
automorphisms of the circuit. To be precise, we can think of a property of
(directed) graphs on $n$ vertices as being given by a Boolean function $f:
\{0,1\}^{n^2} \rightarrow \{0,1\}$ which takes as input the $n^2$ potential
edges and outputs $1$ if the property is satisfied by the graph given on the
input. The function $f$ is \emph{graph-symmetric} in the sense that any permutation of
its inputs induced by a permutation of the $n$ vertices leaves $f$ unchanged
(note this is not the same thing as requiring that $f$ is invariant under all
permutations of its inputs). A symmetric circuit is a circuit computing a
Boolean function of $n^2$ inputs where each permutation of $[n]$ can be extended
to an automorphism of the circuit. Such a circuit necessarily computes an
invariant function.

The results of~\cite{AndersonD17} show a tight relationship between the decdability of
a graph property by means of a polynomial-size family of symmetric circuits and
its definability in extensions of fixed-point logic. In particular, they show
that a property is definable in fixed-point logic with counting ($\FPC$) if, and
only if, it is decidable by a polynomially-uniform family of symmetric circuits
using Boolean and majority gates (or, equivalently, threshold gates). Without
the symmetry restriction, polynomial-size families of circuits have the same
computational power whether we restrict ourselves to Boolean gates only or allow
richer bases such as threshold gates. One consequence of the results
in~\cite{AD14} is that, with the requirement of symmetry, these different bases
do yield different expressive power, suggesting some lack of robustness to the
model. One of the results in the present paper shows that the model with
majority circuits is robust in the sense that it is able to simulate, by means
of symmetric circuits, any symmetric circuit composed of gates computing Boolean
functions that are themselves symmetric (in the sense that they are invariant
under \emph{all} permutations of their inputs. Thus, any property decidable by a
polynomially-uniform family of such circuits is definable in $\FPC$.

This motivates the study of symmetric circuit families that involve gates which
themselves may compute non-symmetric functions. In particular, motivated by the
study of rank logic~\cite{DGHL09,GP15}, we consider gates that are
\emph{matrix-symmetric}. These are gates computing a Boolean function $f:
\{0,1\}^{A \times B} \rightarrow \{0,1\}$ which is invariant under arbitrary
permutations of $A$ and $B$. A typical example is the function computing the
rank of an $A \times B$ $0$-$1$ matrix. We show that the support theorem of
Anderson and Dawar can be shown to hold, even in the presence of
matrix-symmetric gates. This allows us to show that definability in the logic
fixed-point with rank ($\FPR$) is the same as decidability by
polynomially-uniform families of symmetric circuits with rank gates.

% Another extension of $\FPC$ that has received much attention in the descriptive
% complexity literature is the class $\CPTC$ of \emph{choiceless polynomial time
%   with counting}. The work of~\cite{AD14} leaves open the question of obtaining
% a circuit characterization of this class and it was suggested there that this
% would require a relaxation of the notion of symmetry used in defining symmetric
% circuits. In particular, while $\CPTC$ is based on a machine model designed to
% preserve symmetry by forbidding choices, the polynomial time and space
% restrictions apply int he context of the symmetries fo a \emph{single input
%   structure}. In this paper, we consider a relaxation of the notion of symmetric
% circuit intended to capture exactly this intuition. However, we are able to show
% that this does not yield a characterization of $\CPTC$. In particular, we show
% that no family in this class of relaxed symmetric circuits can distinguish odd
% from even Cai-F\"urer-Immerman graphs, something we know is do-able in
% $\CPTC$~\cite{DRR08}.

\chapter{Background}
In this section we present basic background on logics, first-order structures,
and symmetric circuits.

\section{Logic and Structures}
A \emph{relational vocabulary} (usually denoted by $\tau$) is a finite sequence
of relation symbols $R_1, \ldots, R_k$, such that for all $i \in [k]$, $R_i$ has
arity of $r_i \in \nats$.

A $\tau$-structure $\mathcal{A}$ is a tuple $\langle U, R^{\mathcal{A}}_1 ,
\ldots , R^{\mathcal{A}}_k \rangle$, where $U$ is a set and called the
\emph{universe} of $\mathcal{A}$ and for all $i \in [k]$ the relations
$R^{\mathcal{A}}_i \subseteq U^{r_i}$. The elements of $U$ are called the
elements of $\mathcal{A}$. A \emph{multi-sorted} structure is one whose universe
is a disjoint union, where each set in the union is called a \emph{sort}. The
size of a structure is the cardinality of its universe.

All structures in this paper are finite.

\section{First Order, Fixed-Point, Counting}

Let $\FO(\tau)$ denote \emph{first order logic} with respect to the vocabulary
$\tau$. \emph{Fixed-Point Logic} is the extension of $\FO(\tau)$ with an
inflationary fixed-point operator (see \cite{} for details). For logic
$\mathcal{L}$, a formula $\phi \in \mathcal{L}(\tau)$, $x$ a $k$-length tuple of
variables, $\mathcal{A}$ a (finite) structure, and $a \in U^k$, we have that
$\phi (x)$ denotes that the variables in $x$ are free in $\phi$ and we write
$\mathcal{A} \models_{\mathcal{L}} \phi[a]$ iff the the formula $\phi$ is true
in $\mathcal{A}$ with respect to $\mathcal{L}$ with the variables $x$ assigned
to $a$. The context usually makes the logic obvious, and in that case we usually
drop the subscript.

We also recall the definition of \emph{fixed-point logic with counting} $\FPC$,
an extension $\FP$ with counting quantifiers. For a structure $\mathcal{A}$,
Formulas of this logic are evaluated over a structure $\mathcal{A}$ by extending
that structure to $\mathcal{A}^\leq$, a two sorted extension of $\mathcal{A}$ by
an ordered number sort $\{1, \ldots, \vert U\vert \}$. Variables are taken to be
either point variables, ranging over $U$ or number variables ranging over
$[\vert U \vert]$. The counting quantifiers are of the form $\#_x$, where $x$ is
either a point or number variable and for $\phi(x) \in \FPC(\tau)$, $\#_x
\phi(x)$ is a term denoting the element of the number sort corresponding to
$\vert \{a \in \mathcal{A} | \mathcal{A} \models \phi[a]\}\vert$. Note
additionally that $k$-tuples of number variables can be used to encode numbers
in $\FPC$ of size up to $\vert U\vert^k$. For more details on $\FPC$ please see
\cite{}.

\section{Rank Logic}
This still needs to be filled in

\section{Symmetric Circuits}
This still needs to be filled in

\chapter{Symmetric Circuits}
In this section we discuss symmetric circuits, their limitations and extensions
of this model.

\section{Symmetric Circuits and Limitations}
Symmetric circuits, as developed by Anderson and Dawar, are, importantly defined
over (complete) Boolean basis consisting of symmetric functions. The following
result shows that once a majority gate has been included in this basis any
symmetric function can be computed using polynomial-size symmetric circuits, and
as such no additional symmetric function added to the Basis improves the power
of the model.

Recall from Anderson and Dawar \cite{AndersonD17} we have that $\mathbb{B}_{\std} = \{
\neg , \wedge , \lor \}$ and $\mathbb{B}_\maj = \{ \maj \} \cup \mathbb{B}$.

Let $F: \{0,1\}^* \rightarrow \{0,1\}$ be a symmetric Boolean function. Since
$F$ is symmetric we note that for a fixed size input the output of $F$ is
entirely determined by the number of 1's in its input. Then let
$c_{F}:\mathbb{N} \rightarrow 2^{\mathbb{N}}$ define a function where $c_{F}(n)$
is the set of all $m \leq n$ such that for all $\vec{x} \in \{ 0,1 \}^n$ with
$m$ 1's we have $F (\vec{x}) = 1$. Clearly any symmetric Boolean function $F$ is
entirely determined by $c_{F}$.

The above proposition has a straight forward application to circuit
characterisations.

\begin{thm}
  Let $F = \{F_i : i \in I \}$ be a family of symmetric Boolean functions where
  $F_i: \{0,1\}^* \rightarrow \{ 0,1 \}$.

  Let $(C_n)_{n \in \mathbb{N}}$ be family of symmetric circuits over the
  Boolean basis $\mathbb{B} \cup F$, where $C_n$ is a circuit on structures of
  size $n$, and the size of each circuit in the family is bounded by some
  function $f(n)$. Then there exists a polynomial $q(n)$ and a family of
  symmetric circuits $(C_n')_{n \in \mathbb{N}}$ over $\mathbb{B}_\maj$, where
  $C_n'$ is a circuit on structures of size $n$ and $\vert C_n' \vert \leq
  q(f(n))$.
\end{thm}

\section{G-Symmetric Functions}
In order to extend the notion of a circuit by extending the basis we need to
consider functions that are not symmetric in the full sense of the term but have
some weakened notion of symmetry.

\begin{definition}
  Let $f: \{0,1\}^X \rightarrow \{0,1\}$ be a function. Let $G \subseteq \sym_X$
  be a subgroup. We say that $f$ is \emph{$G$-symmetric} if for all $\sigma \in
  G$, $x \in X$, $f(\sigma x) = f(x)$.
\end{definition}

We can also introduce similar notions in the case of a two dimensional Boolean
string.

\begin{definition}
  Let $f: \{0,1\}^{A \times B} \rightarrow \{0,1\}$ be a function. Let $G
  \subseteq \sym_{A \times B}$ be a subgroup. We say that $f$ is
  \emph{matrix-symmetric} if it is $\sym_A \times \sym_B$-symmetric and we say
  that $f$ is \emph{graph-symmetric} if $A = B$ and it is $\sym_A$-symmetric
  (with $\sym_A$ acting on both $A$ and $B$).
\end{definition}

So if we think of $f: \{0,1\}^{A \times B} \rightarrow \{0,1\}$ as representing
a matrix, then matrix-symmetry corresponds to the property of the function being
invariant under row-column permutations. If, instead, we have that $A = B$ and
we then think of $f$ as encoding the adjacency matrix of a graph, then
graph-symmetry corresponds to the function being constant on isomorphism
classes.

Importantly, many natural functions of interest are matrix symmetric. For
example, the function that computes the rank of the matrix over $\mathbb{F}_2$.
or a thresholded rank function, for example the rank of the matrix over
$\mathbb{F}_p$ being larger then $r$, for some particular $(p, r) \in
\mathbb{N}$.

\section{Symmetric Circuits with Matrix-Symmetric Gates}

In this section we develop the corresponding notion of a symmetric circuit, ones
which allow for the inclusion of matrix-symmetric gates. In particular, these
circuits are defined using two sets of Boolean functions. The first, denoted by
$\mathbb{B}_\sym$, is taken to be the symmetric Boolean basis (either
$\mathbb{B}_\std$ or $\mathbb{B}_\maj$). The second, denoted by
$\mathbb{B}_\matsym$ is taken to be the matrix-symmetric basis, consisting of a
collection of matrix-symmetric functions.

\begin{definition}[Circuits on Structures]
  For $\mathbb{B}_{\sym}$ a basis of Boolean symmetric functions,
  $\mathbb{B}_{\matsym}$ a basis of Boolean matrix-symmetric functions and
  $\tau$ a set of relation symbols, we define a $(\mathbb{B}_\sym,
  \mathbb{B}_\sym, \tau)$-circuit $C_n$ computing a $q$-ary query $Q$ is a
  structure $\langle G, W, \Omega, \Sigma, \Lambda, L\rangle$.
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G$ is called the set of gates of $C_n$ and $\vert C_n \vert := \vert G
    \vert$.
  \item $W \subseteq G \times G$, where $W$ is called the wires of the circuit.
    $(G,W)$ must be a directed acyclic graph. For $g \in G$ we $H_g := \{ h \in
    C_n : W(h,g)\}$ be the set of children of $g$.
  \item $\Omega$ is an injective function from $[n]^q$ to $G$. The gates in the
    image of $\Omega$ are called the output gates. When $q = 0$, $\Omega$ is a
    constant function mapping to a single output gate.
  \item $\Sigma$ is a function from $G$ to $\mathbb{B}_\sym \uplus
    \mathbb{B}_\matstab \uplus \tau \uplus \{0,1\} $ which maps input gates to
    $\tau \uplus \{0,1\}$ and where $\Sigma^{-1} (0) \leq 1$ $\Sigma^{-1} (1)
    \leq 1$ and the internal gates get mapped into $\mathbb{B}_\sym \uplus
    \mathbb{B}_\matstab$. Gates mapped to $\tau$ are called relational gates and
    gates mapped to 1 or 0 are called constant gates.
  \item $\Lambda$ is a sequence of injective functions $(\Lambda_R)_{R \in
      \tau}$ where for each $R \in \tau$, $\Lambda_R$ maps each relational gate
    $g$ with $R = \Sigma (g)$ to the tuple $\Lambda_R (g) \in [n]^r$, where $r$
    is the arity of the symbol $R$. When no ambiguity arises we write $\Lambda
    (g)$ for $\Lambda_R (g)$.
  \item $L$ assigns labels to the gates (and their inputs) in $C_n$. Let $g \in
    G$ be an internal gate and $H$ be the gates input to $g$. Then $L(g) : A
    \times B \rightarrow H$, where $L(g)$ must be surjective and $A, B \subseteq
    [n]$. We call this the \emph{matrix labelling} of $g$.
  \end{itemize}
\end{definition}

Note the additional requirement that each gate have an associated labelling in
order to facilitate the evaluation of the gate.

Given some finite $\tau$-structure $\mathcal{A}$ of size $n$ and some bijection
$\gamma: U \rightarrow [n]$, the evaluation of some $(\mathbb{B}_\sym,
\mathbb{B}_\sym, \tau)$-circuit) circuit $C$ proceeds by recursively evaluating
gates. Here the evaluation of the gate $g$ is denoted by $C[\gamma
\mathcal{A}](g)$. If $g$ is anything but a matrix-symmetric gate, then the
evaluation is as per the symmetric circuits of Anderson and Dawar \cite{}. If
$g$ is matrix-symmetric then the evaluation is given by applying the Boolean
operation $\Sigma(g)$ to the matrix $L^{\gamma}:\dom(L(g)) \rightarrow \{0,1\}$
defined by $L^{\gamma}(i,j) = C[\gamma \mathcal{A}](L(g)(i,j))$.

\begin{definition}[Invariant Circuit\cite{AndersonD17}]
  Let $C_n$ be a $(\mathbb{B}_\sym, \mathbb{B}_\sym, \tau)$-circuit, computing
  some $q$-ary query. We say $C_n$ is \emph{invariant} if for every
  $\tau$-structure $\mathcal{A}$ of size $n$, $a \in \mathcal{A}^q$, and
  bijections $\gamma_1, \gamma_2: U \rightarrow [n]$ we have that $C[\gamma_1
  \mathcal{A}](\Omega (\gamma_1 a)) = C[\gamma_2 \mathcal{A}](\Omega (\gamma_2
  a))$.
\end{definition}

\begin{definition}[Automorphism]
  let $C = \langle G, W, \Omega, \Sigma, \Lambda, L\rangle$ be a
  $(\mathbb{B}_\sym, \mathbb{B}_\matsym, \tau)$-circuit computing at $q$-ary
  query on structures of size $n$. Let $\sigma \in \sym_n$ and $\pi: G
  \rightarrow G$ be a bijection such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item for all gates $g, h \in G$, $W(g,h)$ iff $W(\pi g, \pi h)$,
  \item for all output tuples $x \in [n]^q$, $\pi \Omega (x) = \Omega (\sigma
    x)$,
  \item for all gates $g \in G$, $\Sigma (g) = \Sigma (\pi g)$,
  \item for each relational gate $g \in G$, $\sigma \Lambda (g) = \Lambda (\pi
    g)$, and
  \item for each internal gate $g$ if $\Sigma (g) \in \mathbb(B)_\matsym$ then
    we have that $L(\pi g) \sim \pi \cdot L(g)$.
  \end{itemize}

  We call $\pi$ an \emph{automorphism} of $C$, and we say that $\sigma$
  \emph{induces the automorphism} $\pi$. The group of automorphisms of $C$ is
  called $\aut_n (C)$.
\end{definition}

\begin{definition}[Symmetric\cite{AndersonD17}]
  A circuit $C$ on structures of size $n$ is called \emph{symmetric} if every
  $\sigma \in \sym_n$ induces an automorphism on $C$.
\end{definition}

It follows for any symmetric circuit $C_n$ there is a mapping from $\sym_n$ to
$\aut_n(C)$, from $ \sigma$ to the induced automorphism. This homomorphism is
injective so long as a single element of $[n]$ appears in the in the label of
some input gate of $C$ (as then all elements appear by
symmetry)\cite{AndersonD17}. In this paper we always assume that there is always
one such element as otherwise all inputs are constant, and so the circuit just
computes a constant function. In order to assure this homomorphism is surjective
Anderson and Dawar \cite{AndersonD17} introduce the notion of a \emph{rigid}
circuit.

\begin{definition}[Rigidity]
  Let C be a $(\mathbb{B}_\sym, \mathbb{B}_\sym, \tau)$-circuit, where $C =
  \langle G, W, \Omega, \Sigma, \Lambda, L\rangle$. Say that $C$ is rigid if
  there are no internal gates $g, g' \in G$ such that $\Sigma(g) = \Sigma (g')$,
  $\Omega^{-1}(g) = \Omega^{-1}(g')$, and for every $g'' \in G$, $W(g'', g')$
  and $W(g,g'')$ iff $W(g', g'')$.
\end{definition}

Another property which simplifies our analysis is the property of having
\emph{bijective labels}.

We prove in a later section that a circuit may be transformed in polynomial time
into an equivalent circuit that is both rigid and has bijective labelings. Hence
we may assume these two properties without a loss of generality.

With the assumption of rigidity in place, we abuse notation for permutations and
let $\sigma \in \sym_n$ also denote the induced automorphism.

We finally want to define symmetric circuits with rank gates.

\begin{definition}
  Let $\mathbb{B}_{\matsym}$ consist of functions of the form $f_{p,r}:
  \{0,1\}^{A \times B} \rightarrow \{0,1\}$ for $A, B$ initial segments of
  $\mathbb{N}$ and $f, p \in \mathbb{N}$, with $p$ prime, and such that for
  $f_{p,r}(M) = 1$ iff the matrix $M: A \times B \rightarrow \{0,1\}$ has rank
  at least $r$ over the field $\mathbb{F}_p$. A \emph{symmetric circuit with
    rank gates} is a $(\mathbb{B}_\maj, \mathbb{B}_\matsym, \tau)$-circuit, for some
  vocabulary $\tau$.
\end{definition}

The natural restriction to consider on families of circuits is uniformity.

\begin{definition}
  Let $(C_n)_{n \in \mathbb{N}}$ be a family of Boolean circuits. We say that
  $(C_n)_{n \in \mathbb{N}}$ is \emph{$P$-uniform} if the mapping $n \mapsto C_n$ is
  computable in polynomial time.
\end{definition}

We are now ready to state what will be the main theorem of this paper.

\begin{thm}[Main Theorem]
 A graph property is decidable by a $P$-uniform family of symmetric circuits
 with rank gates if, and only if, it is definable by an FPR sentence.
\end{thm}

% \section{Circuits to FPR}

% In this section I discuss how to determine the rank a matrix labelling at some
% gate $g$. This is a rough copy written for the purpose of discussion. In this
% section we fix some circuit $C_n$ and some rank gate $g$ in $C_n$ with a set
% of children $H$ and matrix labelling $L: [a] \times [b] \rightarrow H$ (which
% we assume WLOG is a bijection).

\chapter{Symmetry and Support}
In order to analyse families of symmetric circuits we will need to develop some
theory for analysing symmetries. In this section we define the terms needed
(e.g. the notion of a support) and prove basic results. We then prove the first
important theorem in this paper, the support theorem, using these tools. It is
worth noting that we prove the support theorem for arbitrary symmetric circuits
with matrix-symmetric gates, not just for rank gates.

\section{Supports and Supporting Partitions}
In this section we formally define a support and the related notion of a
supporting partition.

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$ and let $S \subseteq [n]$. Then $S$ is a
  \emph{support} for $G$ if $\stab(S) \subset G$.
\end{definition}

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$ and $\mathcal{P}$ be a partition of $[n]$.
  Then $\mathcal{P}$ is a \emph{supporting partition} for $G$ if
  $\stab(\mathcal{P}) \subseteq [n]$.
\end{definition}

Notice that if $\mathcal{P}$ is a supporting partition for $G$ and $P \in
\mathcal{P}$ then $\stab([n] \setminus P) \subseteq \stab(\mathcal{P}) \subset
G$, i.e. $[n] \setminus P$ is a support for $G$.

These supporting partitions admits a natural partial order in terms of
coarseness. If $\mathcal{P}, \mathcal{P}'$ we say that $\mathcal{P}'$ is as
course as $\mathcal{P}$ (denoted by $\mathcal{P}' \supseteq \mathcal{P}$) if for
all $x \in \mathcal{P}$ there exists $y \in \mathcal{P}'$ such that $x \subseteq
y$ \cite{AndersonD17}.

Anderson and Dawar \cite{AndersonD17} show that for any permutation group $G
\subseteq \sym_n$ a unique coarsest partition can always be found. We call this
partition the \emph{canonical supporting partition}, and denote it by $\SP (G)$.
This canonical supporting partition, along with the above observation about
supports, will allow us to construct for each group a canonical support.

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$. Let $\| SP(G) \| = \min \{\vert [n]
  \setminus P \vert : P \in SP(G) \}$. We say that $G$ has small support if $\|
  SP(G) \| < \frac{n}{2}$.
\end{definition}

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$ such that $G$ has small support. Let
  $\consp(G) = [n] \setminus P$, where $P$ is the maximal element of $\SP(G)$.
  We call $\consp(G)$ the canonical support of $G$.
\end{definition}

\begin{remark}
  Include a remark about the importance of asserting the size consideration. Use
  the alternating group as an example.
\end{remark}
